### 2021-02-23 paper1/paper1.tex
Rearrange things a lot

% ============= content ====================
\chapter{Paper 1: Tuning Fairness by Balancing Target Labels}
\textsc{Authors}:\;\;Thomas Kehrenberg$^1$, Zexun Chen$^{1,2}$, and Novi Quadrianto$^1$ \\
\textsc{Affiliations}:\\
$^1$ Predictive Analytics Lab (PAL), University of Sussex, Brighton, UK\\
$^2$ BioComplex Laboratory, University of Exeter, Exeter, UK \\
\textsc{Journal}:\;\;\textit{Frontiers in Artificial Intelligence}, Volume 3 \\
\textsc{DOI}:\;\;10.3389/frai.2020.00033 \\
\textsc{Note}:\;\;The appendix has been included as section~\ref{sec:tuning-appendix}.

@@

  For $\hat{r}_a > r_a$, the roles of $\mathcal{C}^0$ and $\mathcal{C}^1$ are reversed
  and thus, the signs in the equation are inverted:
  \begin{align}
    \left|\mathcal{A}^\mathit{optimal}\right| = N  - (\left|\mathcal{Y}^1\right| - \left|\mathcal{C}^1\right|)~.
  \end{align}
  This proves the claim.
\end{proof}
\begin{corollary}
  Given a dataset that consists of two subsets $\mathcal{S}_0$
  and $\mathcal{S}_1$ ($\mathcal{D} = \mathcal{S}_0 \cup \mathcal{S}_1$)
  where $p$ is the ratio of $|\mathcal{S}_0|$ to $|\mathcal{D}|$
  and given corresponding acceptance rates $r^0_a$ and $r^1_a$
  and predictions with target rates $\hat{r}^0_a$ and $\hat{r}^1_a$,
  the accuracy is limited by
  \begin{align}
    \mathit{acc} \leq 1 - p\cdot\left| \hat{r}^0_a -r^0_a \right| - (1-p)\cdot\left| \hat{r}^1_a -r^1_a \right|~.
  \end{align}
\end{corollary}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{./figures/plot.pdf}
  \caption{Achievable accuracy for different target values.}%
  \label{fig:accvstarget}
\end{figure}
\begin{example}
  We consider the case where $S_0$
  (which could for example be all data points for female individuals) makes up 30\% of the dataset; so $p = 0.3$.
  Further, we say that for $S_0$ we have an acceptance rate of 10\% ($r^0_a = 0.1$)
  and for $S_1$, 50\% ($r^1_a = 0.5$).
  If we then set both target rates to the same value $t$ ($\hat{r}^0_a=\hat{r}^1_a=t$), with $t = 0.3$,
  then the highest accuracy that can be achieved is $0.8$ or 80\%.

@@

\subsection{Illustration of restrictions on PR}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.3\textwidth]{figures/label_shift.pdf}
  \caption{%
    Illustration of demographic parity with target labels.
    In the situation in the upper part, $P(\bar{y}=1|y=1)$ cannot be set to 1,
    because there are more samples with $y=1$ than there are $\bar{y}=1$.
    In the situation in the lower part, $P(\bar{y}=1|y=1)$ can be set to 1.
  }%
  \label{fig:shift}
\end{figure}%
We start by setting a target rate $r_t$:
\begin{align}
  P(\bar{y}=1|s=0) \overset{!}{=} r_t \quad\text{and}\quad P(\bar{y}=1|s=1) \overset{!}{=} r_t
\end{align}

This leads us to the following constraint for $s\prime\in\{0, 1\}$:
\begin{align}
  r_t &= P(\bar{y}=1|s=s\prime)\nonumber\\
  &= \sum\limits_y P(\bar{y}=1|y,s=s\prime) P(y|s=s\prime)
\end{align}

For $P(y|s=s\prime)$ we will put in the value at which we want our constraint to hold.
We denote $P(y=1|s=j)$ as the base rate $r_b^j$ which we estimate from the training set.
Plugging this in, we are left with
\begin{align}
  r_t = \,\,&P(\bar{y}=1|y=0,s=0) \cdot (1-r_b^0) \nonumber\\
  +\, &P(\bar{y}=1|y=1,s=0) \cdot r_b^0 \\
  r_t = \,\,&P(\bar{y}=1|y=0,s=1) \cdot (1-r_b^1)\nonumber\\
  +\, &P(\bar{y}=1|y=1,s=1) \cdot r_b^1~.
\end{align}
This is a system of linear equations with two equations and four free variables.
There is thus still considerable freedom in how we want our constraint to be realized.
The freedom that we have here concerns how strongly the accuracy will be affected.

@@

We solve for $P(\bar{y}=1|y=0,s=j)$:
\begin{align}
  &P(\bar{y}=1|y=0,s=j)\nonumber\\
  =\,\, &\frac{r_b^j}{1-r_b^j} \left(\frac{r_t}{r_b^j} - P(\bar{y}=1|y=1,s=j)\right)~.
\end{align}
However, we can set $P(\bar{y}=1|y=0,s=j)$ to 0
only if that does not imply $P(\bar{y}=1|y=1,s=j)$ will be greater than 1.
This would happen if $\nicefrac{r_t}{r_b^j}$ were greater than 1.

Figure~\ref{fig:shift} illustrates this.
In the upper part of the figure we have $\nicefrac{r_t}{r_b^j}$ less than 1.
In this case

>>>>

% ============= content ====================
\textsc{Authors}:\;\;Thomas Kehrenberg$^1$, Zexun Chen$^{1,2}$, and Novi Quadrianto$^1$ \\
\textsc{Affiliations}:\\
$^1$ Predictive Analytics Lab (PAL), University of Sussex, Brighton, UK\\
$^2$ BioComplex Laboratory, University of Exeter, Exeter, UK \\
\textsc{Journal}:\;\;\textit{Frontiers in Artificial Intelligence}, Volume 3 \\
\textsc{DOI}:\;\;10.3389/frai.2020.00033 \\
\textsc{Note}:\;\;The appendix has been included as section~\ref{sec:tuning-appendix}.

@@

  For $\hat{r}_a > r_a$, the roles of $\mathcal{C}^0$ and $\mathcal{C}^1$ are reversed
  and thus, the signs in the equation are inverted:
  \begin{align}
    \left|\mathcal{A}^\mathit{optimal}\right| = N  - (\left|\mathcal{Y}^1\right| - \left|\mathcal{C}^1\right|)~.
  \end{align}
  This proves the claim.
\end{proof}
\begin{corollary}
  Given a dataset that consists of two subsets $\mathcal{S}_0$
  and $\mathcal{S}_1$ ($\mathcal{D} = \mathcal{S}_0 \cup \mathcal{S}_1$)
  where $p$ is the ratio of $|\mathcal{S}_0|$ to $|\mathcal{D}|$
  and given corresponding acceptance rates $r^0_a$ and $r^1_a$
  and predictions with target rates $\hat{r}^0_a$ and $\hat{r}^1_a$,
  the accuracy is limited by
  \begin{align}
    \mathit{acc} \leq 1 - p\cdot\left| \hat{r}^0_a -r^0_a \right| - (1-p)\cdot\left| \hat{r}^1_a -r^1_a \right|~.
  \end{align}
\end{corollary}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.6\textwidth]{./figures/plot.pdf}
  \caption{Achievable accuracy for different target values.}%
  \label{fig:accvstarget}
\end{figure}
\begin{example}
  We consider the case where $S_0$
  (which could for example be all data points for female individuals) makes up 30\% of the dataset; so $p = 0.3$.
  Further, we say that for $S_0$ we have an acceptance rate of 10\% ($r^0_a = 0.1$)
  and for $S_1$, 50\% ($r^1_a = 0.5$).
  If we then set both target rates to the same value $t$ ($\hat{r}^0_a=\hat{r}^1_a=t$), with $t = 0.3$,
  then the highest accuracy that can be achieved is $0.8$ or 80\%.

@@

\subsection{Illustration of restrictions on PR}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.6\textwidth]{figures/label_shift.pdf}
  \caption{%
    Illustration of demographic parity with target labels.
    In the situation in the upper part, $P(\bar{y}=1|y=1)$ cannot be set to 1,
    because there are more samples with $y=1$ than there are $\bar{y}=1$.
    In the situation in the lower part, $P(\bar{y}=1|y=1)$ can be set to 1.
  }%
  \label{fig:shift}
\end{figure}%
We start by setting a target rate $r_t$:
\begin{align}
  P(\bar{y}=1|s=0) \overset{!}{=} r_t \quad\text{and}\quad P(\bar{y}=1|s=1) \overset{!}{=} r_t
\end{align}

This leads us to the following constraint for $s\prime\in\{0, 1\}$:
\begin{align}
  r_t = P(\bar{y}=1|s=s\prime)= \sum\limits_y P(\bar{y}=1|y,s=s\prime) P(y|s=s\prime)
\end{align}

For $P(y|s=s\prime)$ we will put in the value at which we want our constraint to hold.
We use $r_b^j$ to denote the base rate $P(y=1|s=j)$ which we estimate from the training set.
Plugging this in, we are left with
\begin{align}
  r_t &= P(\bar{y}=1|y=0,s=0) \cdot (1-r_b^0) +P(\bar{y}=1|y=1,s=0) \cdot r_b^0 \\
  r_t &= P(\bar{y}=1|y=0,s=1) \cdot (1-r_b^1) +P(\bar{y}=1|y=1,s=1) \cdot r_b^1~.
\end{align}
This is a system of linear equations with two equations and four free variables.
There is thus still considerable freedom in how we want our constraint to be realized.
The freedom that we have here concerns how strongly the accuracy will be affected.

@@

We solve for $P(\bar{y}=1|y=0,s=j)$:
\begin{align}
  P(\bar{y}=1|y=0,s=j) = \frac{r_b^j}{1-r_b^j} \left(\frac{r_t}{r_b^j} - P(\bar{y}=1|y=1,s=j)\right)~.
\end{align}
However, we can set $P(\bar{y}=1|y=0,s=j)$ to 0
only if that does not imply $P(\bar{y}=1|y=1,s=j)$ will be greater than 1.
This would happen if $\nicefrac{r_t}{r_b^j}$ were greater than 1.

Figure~\ref{fig:shift} illustrates this.
In the upper part of the figure, we have $\nicefrac{r_t}{r_b^j}$ less than 1.
This means, the target positive rate is less than the base positive rate,
and implies that the positive rate has to be lowered somehow.
This is accomplished by mapping some of the $y=1$ samples to $\bar{y}=0$.
Thus, $P(\bar{y}=1|y=1,s=j)$ is less than 1, and $P(\bar{y}=0|y=1,s=j)$ greater than 0.
In the lower part of the figure, we have the opposite case; essentially, $\bar{y}$ and $y$ swap places.
Here, $P(\bar{y}=1|y=1,s=j)=1$ is possible.
